Flight clustering for route detection and CO{\Large{\textsubscript{2}}} emissions

### 2016-10-23 paper/paper.tex
PAPER: add content

@@

% TODO make a table
%%% position grouping
\begin{table}
  \centering
  \begin{tabular}{|c|c|c|}
  \hline
  \textbf{Day in dataset} & \textbf{Clusters} & \textbf{Detected flights} \\
  \hline
  18/09/2016 & 199030 & 1219 \\
  \hline
  19/09/2016 & 170594 & 1103 \\
  \hline
  20/09/2016 & 126370 & 1241 \\
  \hline
  21/09/2016 & 127855 & 1205 \\
  \hline
  22/09/2016 & 127982 & 1308 \\
  \hline
  23/09/2016 & 130662 & 1243 \\
  \hline
  24/09/2016 & 121624 & 874 \\
  \hline
  \end{tabular}
  \caption{Number of clusters and flights per day}
  \label{table}
\end{table}

@@

In order to compute the amount of flights diverging from a standard route per
airline, we needed to associate a an airline to each flight. This is not a
completely trivial task, as airline information is not provided at all by the
data. Every aircraft, hence every flight, has an ICAO number associated to it,
from which it is possible to determine callsigns from the identification
messages in the dataset. In the case of commercial flights, the callsign of an
aircraft contains the ICAO three-letter code of the airline for which the
vehicle is flying for at the moment. Since aircraft are bought by airlines and
hardly transferred from an airline to another, it is safe to assume that only
one callsign is sufficient to determine the airline of an aircraft.
The mapping between three-letter ICAO codes of airlines and their name has been
done with the help of an external dataset.

Of 19860 distinct icaos for which identification messages were available, 15911
had a unique three-letter airline code in their callsigns, and 15691 had a code
actually corresponding to an airline in the dataset. The remaining 4169
icaos without a proper airline code are likely to correspond to private airplanes.

@@

>>>>

Every ADS-B transmitter is assigned a unique 24-bit ICAO number. Since the
transmitters installed on the airplanes are hardly changed, we can safely
consider an aircraft and its ADS-B transmitter as the same entity. Every raw
ADS-B message records the ICAO number of the transmitter, hence the aircraft,
that sent it. This is basically the way we are able to associate information
from the ADS-B messages to individual airplanes.

@@

% TODO make a table
%%% position grouping
\begin{table}
  \centering
  \begin{tabular}{|c|c|c|}
  \hline
  \textbf{Day in dataset} & \textbf{Clusters} & \textbf{Detected flights} \\
  \hline
  18/09/2016 & 199030 & 1219 \\
  \hline
  19/09/2016 & 170594 & 1103 \\
  \hline
  20/09/2016 & 126370 & 1241 \\
  \hline
  21/09/2016 & 127855 & 1205 \\
  \hline
  22/09/2016 & 127982 & 1308 \\
  \hline
  23/09/2016 & 130662 & 1243 \\
  \hline
  24/09/2016 & 121624 & 874 \\
  \hline
  \end{tabular}
  \caption{Amount of clusters and flights per day, as determined by the flight
    detection algorithm}
  \label{table}
\end{table}

@@

In order to compute the amount of flights diverging from a standard route per
airline, we needed to associate a an airline to each flight. This is not a
completely trivial task, as airline information is not provided at all by the
data. Every aircraft, hence every flight, has an ICAO number associated to it,
from which it is possible to determine callsigns from the identification
messages in the dataset. In the case of commercial flights, the callsign of an
aircraft contains the three-letter code of the airline for which the
vehicle is flying for at the moment. Since aircraft are bought by airlines and
hardly transferred from an airline to another, it is safe to assume that only
one callsign is sufficient to determine the airline of an aircraft.
The mapping between three-letter codes of airlines and their name has been
done with the help of an external dataset.

Of 19860 distinct ICAOs for which identification messages were available, 15911
had a unique three-letter airline code in their callsigns, and 15691 had a code
actually corresponding to an airline in the dataset. The remaining 4169
ICAOs without a proper airline code are likely to correspond to private airplanes.

@@

We were able to associate an airline to almost all of the flights the resulted
from the route clustering phase. From this information, we then determined how
many flights, per airline, follow and do not follow the standard route for their
particular pair of airports, as identified by our standard route detection.
We observed that the majority of flights do follow tha standard route for their
particular pair of departure and arrival airports. Table~\ref{airline-table}
shows an excerpt of the data that we were able to collect.

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|}
  \hline
  \textbf{Airline} & \textbf{Normal} & \textbf{Diverging} \\
  \hline
  KLM Royal Dutch Airlines & 203 & 14 \\
  Ryanair & 124 & 20 \\
  Swiss International Air Lines & 794 & 38 \\
  Alitalia & 412 & 39 \\
  British Airways & 12 & 2 \\
  Lufthansa & 1329 & 38 \\
  easyJet & 108 & 15 \\
  Germanwings & 191 & 18 \\
  \hline
  \end{tabular}
  \caption{Excerpt of data relating some major airlines to flights that comply
    or diverge from a standard route}
  \label{airline-table}
\end{table}

From these results we may conclude that, especially for big companies with many
flights associated to them, the standard routes as detected by our algorithm are
really ``standard'', in the sense that there is a significant difference between
flights that follow them and flights that do not.
